### 2021-04-09 README.Rmd

---
output: github_document
---

# README

[![build](https://github.com/tbep-tech/wqtrends-manu/workflows/build/badge.svg)](https://github.com/tbep-tech/wqtrends-manu/actions)
Materials for water quality trends assessment using GAMs
[![DOI](https://zenodo.org/badge/249504727.svg)](https://zenodo.org/badge/latestdoi/249504727)

## Data

* `datprc.RData`: Formatted water quality data from <https://github.com/fawda123/SFbaytrends>

* `mods_chl18.RData` through `mods_chl36.RData` also from <https://github.com/fawda123/SFbaytrends>

## Docs

* Outline <https://tbep-tech.github.io/wqtrends-manu/outline.html>

* Literature <https://tbep-tech.github.io/wqtrends-manu/lit.html>

* Manuscript draft (rendered in Word) <https://tbep-tech.github.io/wqtrends-manu/manu_draft.docx>