### 2021-05-05 supplement.Rmd
5 and 10 year trend comparisons added to supplement

---
title: "Supplement to Multi-scale trend analysis of water quality using error propagation of generalized additive models"
output: 
  bookdown::word_document2:
    reference_docx: my_styles.docx
    number_sections: false
always_allow_html: true
bibliography: refs.bib
author: "Marcus W. Beck (mbeck@tbep.org), Perry de Valpine (pdevalpine@berkeley.edu), Rebecca Murphy (rmurphy@chesapeakebay.net), Ian Wren (ianw@sfei.org), Ariella Chelsky (ariellac@sfei.org), Melissa Foley (melissaf@sfei.org), David B. Senn (davids@sfei.org)"
urlcolor: blue
csl: ecological-applications.csl
link-citations: true
---

```{r setup, echo = F, warning = F, message = F, results = 'hide'}
# figure path, chunk options
knitr::opts_chunk$set(fig.path = 'figs/', warning = F, message = F, echo = F, cache = F, dev.args = list(family = 'serif'), dpi = 300, warning = F,
  fig.process = function(x) {
  x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
  if (file.rename(x, x2)) x2 else x
  })

# libraries
library(Jabbrev)
library(tidyverse)
library(wqtrends)
library(flextable)
library(officer)
library(ggmap)
library(sf)
library(ggsn)
library(USAboundaries)
library(rworldmap)
library(hrbrthemes)
library(patchwork)
library(lubridate)
library(colorspace)

data(locs)
data(modprf)
data(seastrnd)
data(seastrnd2)
data(cmptrnd05)
data(cmptrnd15)
data(modstr)
data(modslog_chl18)
data(modslog_chl24)
data(modslog_chl27)
data(modslog_chl32)
data(modslog_chl34)

source('R/funcs.R')

locs <- locs %>% 
  rename(station = Station)

# # extract bib entries from online
# bib_scrp('manu_draft.Rmd', 'refs.bib')
```

`r paste('Last build', Sys.time())`

```{r echo = F, cache = F, eval = F}
spelling::spell_check_files('manu_draft.Rmd')
```

# Comments on the `select` argument in `mgcv

 When `select = TRUE` is included, the comparison between models S and SY changes.  This option tells `mgcv` to penalize the coefficient of the linear terms in the spline.  This would be appropriate if `cont_year` was an explanatory variable subject to variable selection, but it is irrelevant if including both a linear and spline term for `cont_year`.  If `select = TRUE` is used, models S and SY would still be effectively equivalent, but AIC selection would suggest that one model is superior.  This result would be an artifact of the choice in model SY to include a linear trend in `cont_year` both as a separate term and as part of the spline, with the latter subject to penalization.

# Back-transformation of model results

Model results were back-transformed from log$_{10}$-space to aid in the interpretation of trends [@Bradu70;@Duan83]. Back-transformation was accomplished using equation \@ref(eq:backtrans) for estimates of mean values and endpoints of confidence intervals from GAM results, such that:

\begin{equation}
E[y] = 10^{\mu + \frac{\log(10) \sigma^2}{2}}
(\#eq:backtrans)
\end{equation}

where the back-transformed, expected value of the response variable $y$ (chl-a) is a function of the predicted value $\mu$ (mean or confidence interval endpoint) in log$_{10}$-space and a dispersion estimate $\sigma^2$ from the model.  The dispersion is the residual variance estimated from the GAM fit.

# Figures

```{r, results = 'hide', message = F}
# get predictions to plot
modprds <- list.files('data', pattern = '^modslog\\_chl', full.names = T) %>% 
  enframe() %>% 
  group_by(value) %>% 
  nest %>% 
  mutate(
    prd = purrr::map(value, function(x){

      load(file = x)

      nm <- basename(x)
      nm <- gsub('\\.RData', '', nm)

      mod <- get(nm) %>% 
        pull(model) %>% 
        deframe()

      prd <- anlz_prd(mod)

      return(prd)

    })
  ) %>% 
  ungroup %>% 
  select(station = value, prd) %>% 
  mutate(
    station = gsub('^data/modslog\\_chl|\\.RData$', '', station)
  ) %>% 
  unnest('prd')

modobs <- rawdat %>% 
  filter(param == 'chl')
ylab <- expression(paste("Chl-a (", mu, "g ", L^-1, ")"))

stas <- unique(modprds$station)

for(sta in stas){

  toplo1 <- modprds %>% 
    filter(station %in% sta)
  toplo2 <- modobs %>% 
    filter(station %in% sta)

  p1 <- ggplot2::ggplot(toplo1, ggplot2::aes(x = date)) + 
    ggplot2::geom_point(data = toplo2, ggplot2::aes(y = value), size = 0.5, colour = 'grey') +
    ggplot2::geom_line(ggplot2::aes(y = value), size = 0.75, alpha = 0.7) + 
    ggplot2::scale_color_viridis_d() + 
    ggplot2::theme_minimal(base_family = 'serif', base_size = 14) + 
    ggplot2::scale_y_log10(limits = c(0.26, 207)) +
    ggplot2::theme(
      legend.position = 'top', 
      legend.title = ggplot2::element_blank(),
      strip.background = ggplot2::element_blank()
    ) + 
    ggplot2::labs(
      y = ylab,
      x = 'Continuous year', 
      subtitle = paste0('Station ', sta, ' continuous and day of year predictions')
    )

  p2 <- ggplot2::ggplot(toplo1, ggplot2::aes(x = doy, group = factor(yr), colour = yr)) + 
    ggplot2::geom_line(ggplot2::aes(y = value), size = 0.5, alpha = 1) + 
    ggplot2::theme_minimal(base_family = 'serif', base_size = 14) + 
    ggplot2::theme(
      legend.position = 'right', 
      legend.title = ggplot2::element_blank(), 
      strip.background = ggplot2::element_blank(), 
      strip.text = ggplot2::element_blank(), 
      axis.title.y = ggplot2::element_blank()
    ) + 
    ggplot2::scale_colour_gradient(low = "white", high = "black") + 
    ggplot2::scale_y_log10(limits = c(0.26, 207)) +
    ggplot2::guides(colour = ggplot2::guide_colourbar(barheight = 10, barwidth = 1)) +
    ggplot2::labs(
      x = "Day of year"
    )
  p <- p1 + p2 + plot_layout(ncol = 2, width = c(1, 0.5))

  # # save as tiff
  # tiff(paste0('figs/prddat', sta, '.tiff'), height = 3.5, width = 9, family = 'serif', units = 'in', res = 600, compression = 'lzw')
  # print(p)
  # dev.off()

  # save as png
  png(paste0('figs/prddat', sta, '.png'), height = 3.5, width = 9, family = 'serif', units = 'in', res = 400)
  print(p)
  dev.off()

}
```

```{r prddat18, fig.cap = "GAM predictions for station 18 for model S. The results show predictions across the time series and predictions by day of year.  Observed data in are shown with the gray points."}
knitr::include_graphics('figs/prddat18.png')
```

```{r prddat21, fig.cap = "GAM predictions for station 21 for model S. The results show predictions across the time series and predictions by day of year.  Observed data in are shown with the gray points."}
knitr::include_graphics('figs/prddat21.png')
```

```{r prddat22, fig.cap = "GAM predictions for station 22 for model S. The results show predictions across the time series and predictions by day of year.  Observed data in are shown with the gray points."}
knitr::include_graphics('figs/prddat22.png')
```

```{r prddat24, fig.cap = "GAM predictions for station 24 for model S. The results show predictions across the time series and predictions by day of year.  Observed data in are shown with the gray points."}
knitr::include_graphics('figs/prddat24.png')
```

```{r prddat27, fig.cap = "GAM predictions for station 27 for model S. The results show predictions across the time series and predictions by day of year.  Observed data in are shown with the gray points."}
knitr::include_graphics('figs/prddat27.png')
```

```{r prddat30, fig.cap = "GAM predictions for station 30 for model S. The results show predictions across the time series and predictions by day of year.  Observed data in are shown with the gray points."}
knitr::include_graphics('figs/prddat30.png')
```

```{r prddat32, fig.cap = "GAM predictions for station 32 for model S. The results show predictions across the time series and predictions by day of year.  Observed data in are shown with the gray points."}
knitr::include_graphics('figs/prddat32.png')
```

```{r prddat34, fig.cap = "GAM predictions for station 34 for model S. The results show predictions across the time series and predictions by day of year.  Observed data in are shown with the gray points."}
knitr::include_graphics('figs/prddat34.png')
```

```{r prddat36, fig.cap = "GAM predictions for station 36 for model S. The results show predictions across the time series and predictions by day of year.  Observed data in are shown with the gray points."}
knitr::include_graphics('figs/prddat36.png')
```

```{r, results = 'hide', message = F}
toplo <- cmptrnd05
p <- ggplot(toplo, aes(y = station, x = yrcoef, linetype = pval, shape = modtyp, colour = modtyp)) + 
  facet_grid(seas ~ yrs, scales = 'free_x') + 
  geom_point(size = 2, position = position_dodge(width = 0.6), alpha = 0.8) + 
  geom_errorbarh(aes(xmin = yrcoeflwr, xmax = yrcoefupr), position = position_dodge(width = 0.6), height = 0) +
  scale_linetype_manual('Trend significance', values = c('dashed', 'solid')) + 
  scale_colour_discrete_qualitative(name = 'Trend model', palette = 'Dark3') +
  scale_shape(name = 'Trend model') + 
  geom_vline(xintercept = 0, linetype = 'dashed') + 
  # scale_x_continuous(limits = c(-1.14, 1.14)) +
  guides(
    color = guide_legend(override.aes = list(linetype = 0))
  ) +
  theme_minimal() + 
  theme(
    strip.background = element_blank(),
    panel.background = element_rect(fill = 'gray98', colour = NA), 
    panel.border = element_blank(), 
    legend.position = 'top', 
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(
    y = 'Station (north to south)', 
    x = expression(paste(log[10], " chl-a change (", italic(mu), "g ", L^-1, yr^-1, ")"))
  )

# save as png
png('figs/trndcmp05.png', height = 6, width = 7.5, family = 'serif', units = 'in', res = 400)
p
dev.off()
```
```{r trndcmp05, fig.cap = 'Trend estimate comparisons for three models applied to seasonal averages of chl-a in 5 year periods at each station. The "OLS raw" trend model is based on an ordinary least squares (OLS) regression fit to the seasonal averages of chl-a from the raw data, the "OLS GAM" trend model is based on an OLS regression fit to the seasonal averages of chl-a from the GAM model (without error propagation), and the "Meta-analysis GAM" trend model is based on a meta-analysis regression fit to the seasonal averages of chl-a from the GAM model.  Values for each model are the log$_{10}$-slope estimates (+/- 95\\% confidence interval) as annual change per year within each season, with line style denoting trend significance.'}
knitr::include_graphics('figs/trndcmp05.png')
```

```{r, results = 'hide', message = F}
toplo <- cmptrnd15
p <- ggplot(toplo, aes(y = station, x = yrcoef, linetype = pval, shape = modtyp, colour = modtyp)) + 
  facet_grid(seas ~ yrs, scales = 'free_x') + 
  geom_point(size = 2, position = position_dodge(width = 0.6), alpha = 0.8) + 
  geom_errorbarh(aes(xmin = yrcoeflwr, xmax = yrcoefupr), position = position_dodge(width = 0.6), height = 0) +
  scale_linetype_manual('Trend significance', values = c('dashed', 'solid')) + 
  scale_colour_discrete_qualitative(name = 'Trend model', palette = 'Dark3') +
  scale_shape(name = 'Trend model') + 
  geom_vline(xintercept = 0, linetype = 'dashed') + 
  # scale_x_continuous(limits = c(-1.14, 1.14)) +
  guides(
    color = guide_legend(override.aes = list(linetype = 0))
  ) +
  theme_minimal() + 
  theme(
    strip.background = element_blank(),
    panel.background = element_rect(fill = 'gray98', colour = NA), 
    panel.border = element_blank(), 
    legend.position = 'top'
  ) +
  labs(
    y = 'Station (north to south)', 
    x = expression(paste(log[10], " chl-a change (", italic(mu), "g ", L^-1, yr^-1, ")"))
  )

# save as png
png('figs/trndcmp15.png', height = 6, width = 7.5, family = 'serif', units = 'in', res = 400)
p
dev.off()
```
```{r trndcmp15, fig.cap = 'Trend estimate comparisons for three models applied to seasonal averages of chl-a in 15 year periods at each station. The "OLS raw" trend model is based on an ordinary least squares (OLS) regression fit to the seasonal averages of chl-a from the raw data, the "OLS GAM" trend model is based on an OLS regression fit to the seasonal averages of chl-a from the GAM model (without error propagation), and the "Meta-analysis GAM" trend model is based on a meta-analysis regression fit to the seasonal averages of chl-a from the GAM model.  Values for each model are the log$_{10}$-slope estimates (+/- 95\\% confidence interval) as annual change per year within each season, with line style denoting trend significance.'}
knitr::include_graphics('figs/trndcmp15.png')
```

# References