### 2021-06-24 revision/review_response.Rmd
more STE revisions

---
output: 
  bookdown::word_document2:
    reference_docx: response_styles.docx
    number_sections: false
always_allow_html: true
urlcolor: blue
link-citations: true
---

```{r setup, echo = F, warning = F, message = F, results = 'hide'}
# figure path, chunk options
knitr::opts_chunk$set(fig.path = 'figs/', warning = F, message = F, echo = F, cache = F, dev.args = list(family = 'serif'), dpi = 300, warning = F)
```

```{r echo = F, cache = F, eval = F}
spelling::spell_check_files('revision-response.Rmd')
```

We sincerely thank the reviewers for providing thoughtful comments on our manuscript.  We have provided a point-by-point response to these comments below.  We are confident that these additions and revisions improve the content of our manuscript for readers of Science of the Total Environment.

## Reviewer 1

This manuscript is a welcome addition in the world of trend analyses. The authors propose combinations of GAM and mixed-effects meta-analysis regression to account for uncertainty. The methodology is sound and results are convincing. My recommendation would be even more enthusiastic if the authors had formatted their methods sections in a more traditional manner. I am of the opinion that models should be presented using clear mathematical equations and not in R code semantics. This comments and a few other can be found in the attached document.

* __Response__: We really appreciate these comments that affirm our opinion on the value of this work for trend analysis of long-term water quality data.  We have changed our formulas and text to a conventional format, as opposed to R code notation.  Briefly, here are the updated equations for the models described in the methods.

\begin{equation}
\textrm{Model S}: y_i \sim\ \beta_0 + f_1\left(cont\_year_i\right) + \epsilon_i
(\#eq:Smod)
\end{equation}

\begin{equation}
\textrm{Model SY:} y_i \sim\ \beta_0 + \beta_1\left(cont\_year_i\right) + f_1\left(cont\_year_i\right) + \epsilon_i
(\#eq:SYmod)
\end{equation}

\begin{equation}
\textrm{Model SYD:} y_i \sim\ \beta_0 + \beta_1\left(cont\_year_i\right) + f_1\left(cont\_year_i\right) + f_2\left(doy_i) + \epsilon_i
(\#eq:SYDmod)
\end{equation}

\begin{equation}
\textrm{Model SYDI:} y_i \sim\ \beta_0 + \beta_1\left(cont\_year_i\right) + f_1\left(cont\_year_i\right) + f_2\left(doy_i) + f_3\left(cont\_year_i, doy_i\right) + \epsilon_i
(\#eq:SYDImod)
\end{equation}

## Reviewer 2

### Overview

This manuscript presents a new application of GAMs approaches to fit irregular time series data (gaps, changing frequency of measurements) to estimate trends. The paper presents a case study using Chl a data from San Francisco Bay.  The manuscript is presented as a general approach with applications to other similar environmental time series data.  A comparison of methods shows that the GAMs plus meta-analysis method provides some different trend conclusions from ordinary least squares and plain GAMs. The paper is generally well written and well-suited for the audience for STOTEN.  I have some clarifications that would help improve/clarify the utility and readability of the manuscript.  These are presented below.

* __Response__: Thank you for your comment on our manuscript.  We have addressed your concerns on the utility and readability of our manuscript below.

### Utility (1):

We applied the wqtrends code in R to a time series data set of X2 (salinity intrusion measure) over approximately in San Francisco Bay.  Some observations:

-the code did not work with the full time series of daily X2 data.  But we were able to make it work using a random sampling of 5 points per month.  Could the reason for this error be explored?  As time series data are concerned, we provided an input of ~100 years of daily data.  A sample result for the trends are shown in the figure below.  I think it would be helpful for readers/users to understand what the practical limitations of the code/algorithm are.

* __Response__: Very rarely do reviewers actually apply the proposed methods on novel datasets as a proof of concept.  For that, we are very appreciative of your efforts in exploring the R package. These applications are important for testing both the method and software.  As such, you exposed an important limitation of our method that was not clearly articulated in the draft.  We provide two responses to this limitation.

     First, the X2 dataset is collected daily, so the GAM smoothing methods proposed in our manuscript for "irregularly spaced or missing" monitoring data is unnecessary.  The modelling approach is designed to provide both a continuous (e.g., daily) estimate of the long-term trend and a measure of uncertainty in the absence of more regularly collected values.  So, this would be an unnecessary applications of GAMs.

     Second, and more important, the practical limitations of applying a GAM with flexible smoothing to nearly 100 years of daily data is beyond the computational power most desktop computers.  This is asking too much for the smoothing splines used in the mgcv package. Given this limitation, we do feel it is important to clearly state the intended use case for these models to avoid confusion in application.  We have added clarification both in the manuscript and the package vignette so users are clear on the type of data that is appropriate for the methods.

     In the manuscript, line xyx: "Approximate monthly or biweekly sampling with coverage of at least a decade is common for many long-term monitoring programs and is the the intended use case for the methods herein."

     In the package vignette, first paragraph (https://tbep-tech.github.io/wqtrends/articles/introduction.html): "These models are appropriate for data typically collected for surface water quality monitoring programs at roughly monthly or biweekly collection intervals, covering at least a decade of observations (e.g., Cloern and Schraga 2016). Daily or continuous monitoring data are not appropriate for these methods, due to computational limitations and a goal of the analysis to estimate long-term, continuous trends from irregular or discontinuous sampling."

### Utility (2):

From figure 7 it seems that there are very different conclusions for Station 36 depending on the method used.  From Figure 8, however, it seems that Station 36 is a bit of an anomaly in this regard, and that other 8 stations in the case study have findings on trends that are similar among the three methods—and that the results are not dramatically different.  Perhaps this should be the basis for added discussion in the final section—when is the added complexity of the GAMS and meta analysis method is appropriate in a real-world setting?  Alternatively, it would be helpful if the authors were to examine another data set from the region that they work with, to review the relative findings of trend significance and what this depends on.

* __Response__: You are correct to note that many of the comparisons had similar results and we agree that it is useful to clarify that our approach is warranted relative to more conventional methods.  If a simpler approach is used, there is no way to know for sure that a false positive or negative result is obtained without applying the proposed methods.  So, we argue that using our approach is still the wiser option in all cases.  Otherwise, a user cannot be certain the correct conclusions are obtained. We have added the following to the discussion:

    Line xyz: "Moreover, use of methods that fully account for uncertainty in sampling design is recommended over more conventional analyses given the certainty of conclusions in the former.  Our examples demonstrated the increased potential for incorrect conclusions at specific monitoring locations and at much larger spatial scales across all stations if simpler trend analyses are used. Even though simpler methods may produce the same result, particularly with frequent sampling and similar effort between years, decision-makers require certainty of conclusions when incorrect results can have expensive consequences."

### Readability/Clarity:

The term meta-analysis seems to be an important part of the paper and is referred to multiple times, but it is never clearly explained in the manuscript.  For example, at line 164.  See the following sentences.

"Third, we used a mixed-effects meta-analysis to estimate trends and test hypotheses about the change in seasonal averages across years. While 166 meta-analysis methods arose from analyses of results from multiple studies, their distinguishing characteristic is propagation of uncertainty (Gasparrini et al., 2012; Sera et al., 2019). Meta-analysis uses response data that includes standard errors (uncertainties) as needed to address our questions."

It is not clear to me what exactly is meant by meta-analysis and what is being tested in the context of the time series data.  An explanation that is understandable by a general audience of potential users of this method would be helpful.  It does seem likely, as that sentence points out, that most readers would be more familiar with meta-analysis methods in the context of aggregating the results of multiple studies.  There are other helpful citations afterwards, but a more explicit connection between the familiar use case and the use case here of combining different seasonal features from a single regression model would have been helpful for reading this work.

* __Response__: Thank you for point out this ambiguity. We agree that a distinction between the more common understanding of "meta-analysis" and the one applied here needs to be made early in the manuscript.  We have added content in the introduction to clarify our methods.

     Line xyz: "Herein, meta-analysis refers to the use of seasonal features and their uncertainty estimates for trend analysis, as compared to meta-analysis for the aggregation of results from multiple studies for analysis of effect sizes or other synthesis results (i.e., Lortie et al. 2014)."